---
title: "R Notebook"
output: html_notebook
Name: Joseph Randolph 

---

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

5.1.1 Prerequisites

```{r}
library(tidyverse)
library(dplyr)
library(nycflights13)
library(ggplot2)
```

>Take careful note of the conflicts message that’s printed when you load the tidyverse. It tells you that dplyr overwrites some functions in base R. If you want to use the base version of these functions after loading dplyr, you’ll need to use their full names: stats::filter() and stats::lag().

```{r}
flights # This data frame contains all 336,776 flights that departed from New York City in 2013. The data comes from the US Bureau of Transportation Statistics.It prints differently because it’s a tibble. it only shows the first few rows and all the columns that fit on one screen
```
>the row of three (or four) letter abbreviations under the column names describe the type of each variable:

>int stands for integers.

>dbl stands for doubles, or real numbers.

>chr stands for character vectors, or strings.

>dttm stands for date-times (a date + a time).

>lgl stands for logical, vectors that contain only TRUE or FALSE.

>fctr stands for factors, which R uses to represent categorical variables with fixed possible values.

>date stands for dates.

```{r}
View(flights) # (To see the whole dataset, you can run View(flights) which will open the dataset in the RStudio viewer). 
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

5.1.3 dplyr basics

>the five key dplyr functions that allow you to solve the vast majority of your data manipulation challenges:

> Pick observations by their values (filter()).
> Reorder the rows (arrange()).
> Pick variables by their names (select()).
> Create new variables with functions of existing variables (mutate()).
> Collapse many values down to a single summary (summarise()).

>These can all be used in conjunction with group_by() which changes the scope of each function from operating on the entire dataset to operating on it group-by-group

5.2 Filter rows with filter()

>filter() allows you to subset observations based on their values. 

~The first argument is the name of the data frame. 

~The second and subsequent arguments are the expressions that filter the data frame. 

~For example, we can select all flights on January 1st with:

> expl filter(flights, month == 1, day == 1)

```{r}
jan1 <- filter(flights, month == 1, day == 1) #dplyr executes the filtering operation and returns a new data frame
jan1
```

5.2.1 Comparisons

>To use filtering effectively, you have to know how to select the observations that you want using the comparison operators. R provides the standard suite: >, >=, <, <=, != (not equal), and == (equal).

When you’re starting out with R, the easiest mistake to make is to use = instead of == when testing for equality. When this happens you’ll get an informative error:

filter(flights, month = 1)
#> Error: `month` (`month = 1`) must not be named, do you need `==`?

There’s another common problem you might encounter when using ==: floating point numbers. These results might surprise you!

sqrt(2) ^ 2 == 2
#> [1] FALSE
1 / 49 * 49 == 1
#> [1] FALSE

>For other types of combinations, you’ll need to use Boolean operators yourself: & is “and”, | is “or”, and ! is “not”

```{r}
nov_dec <-filter(flights, month == 11 | month == 12) #The code finds all flights that departed in November or December
nov_dec
```


!(x & y) is the same as !x | !y, and !(x | y) is the same as !x & !y. For example, if you wanted to find flights that weren’t delayed (on arrival or departure) by more than two hours, you could use either of the following two filters:

```{r}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```

5.2.3 Missing values

Computers use finite precision arithmetic (they obviously can’t store an infinite number of digits!) so remember that every number you see is an approximation. Instead of relying on ==, use near()
```{r}
near(sqrt(2) ^ 2,  2)
```

```{r}
near(1 / 49 * 49, 1)
```


 You can’t write filter(flights, month == 11 | 12), which you might literally translate into “finds all flights that departed in November or December”. Instead it finds all months that equal 11 | 12, an expression that evaluates to TRUE
```{r}
filter(flights, month == 11 | 12)
```
 
 A useful short-hand for this problem is x %in% y. This will select every row where x is one of the values in y. We could use it to rewrite the code above:
```{r}
nov_dec <- filter(flights, month %in% c(11, 12))
nov_dec
```

 !(x & y) is the same as !x | !y, and !(x | y) is the same as !x & !y. For example, if you wanted to find flights that weren’t delayed (on arrival or departure) by more than two hours, you could use either of the following two filters:
```{r}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```
 
5.2.3 Missing values

I'm not undwerstanding this setion about the NA's

One important feature of R that can make comparison tricky are missing values, or NAs (“not availables”). NA represents an unknown value so missing values are “contagious”: almost any operation involving an unknown value will also be unknown.

NA > 5
#> [1] NA
10 == NA
#> [1] NA
NA + 10
#> [1] NA
NA / 2
#> [1] NA

The most confusing result is this one:

NA == NA
#> [1] NA

It’s easiest to understand why this is true with a bit more context:

# Let x be Mary's age. We don't know how old she is.
x <- NA

# Let y be John's age. We don't know how old he is.
y <- NA

# Are John and Mary the same age?
x == y
#> [1] NA
# We don't know!

If you want to determine if a value is missing, use is.na():

is.na(x)
#> [1] TRUE



>5.3 Arrange rows with arrange()



